Smart Model Elimination for Efficient Automated Machine Learning

### 2024-06-16 paper.tex
dataset attributes

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{geometry}
\usepackage{longtable}
\usepackage{graphicx}

@@

\textbf{Ease of Use} & Easy-to-use GUI and API & Python library, requires coding experience & Very user-friendly web interface, detailed API & Simple API & User-friendly API with minimal coding & Python library with fairly intuitive API\\
\hline
\textbf{Supported Algorithms} & Linear, Tree, DNN, Clustering, etc. & Genetic Programming & Linear, Tree, LightGBM, CatBoost, Clustering, etc. (no DNN) & Linear, Tree, Clustering - More limited and lightweight & Linear, LightGBM, Catboost, etc. & Linear, tree, SVM, etc. \\
\hline
\textbf{Feature Engineering} & Basic & Basic & Advanced - includes feature interactions & Limited & Advanced (automatic) & Basic \\
\hline
\textbf{Time Management} & Time Constraints are Settable & Generations and population size & Time constraints are Settable & Optimized for quick results & Settable time constraints & Settable time constraints\\
\hline
\textbf{Ensemble Methods} & Yes (stacking + blending) & Yes (stacking) & Yes (stacking + blending) & No & Yes (stacking) & Yes (stacking) \\
\hline
\textbf{Hyperparameter Tuning} & Automated & Genetic algorithm & Automated and customizable & Automated and efficient tuning & Automated & Automated \\
\hline
\textbf{Model Interpretability} & SHAP, LIME, PDPs, MOJO & POJO, etc. & Basic & SHAP, LIME, PDPs, etc. & Basic (permutation importance) & SHAP, PDPs, etc. & Basic (feature importance) \\
\hline
\textbf{Scalability} & Good (distributed computing) & Limited & Good & Good (resource-efficient) & Good (distributing computing) & Moderate \\
\hline
\textbf{Customization} & High & Moderate & High & Low-Moderate & Moderate & Moderate \\
\hline

@@

Taking a closer look at four AutoML standards, we can observe pros and cons in their theoretical application to medicine and research. FLAML uses blend search hyper-parameter optimization and focuses on lightweight models, making it quick and applicable to repetitive tasks in the medical sphere. It supports imputation techniques for missing values with user specification and has an intuitive API, making it moderately user-friendly for minimally trained healthcare workers. H2O has a wider variety of supported algorithms that may or may not be appropriate for medical diagnosis, making it slower than FLAML on average. In one trial, more than half of FLAML's performances in one minute were better than or equal to H20's performances in one hour. H2O has automatic and multi-faceted methods of dealing with missing values and flexible interfaces in R and Python, making it intuitive for users. MLJAR uses more advanced algorithms such as light gradient boosting and neural networks. It automatically deals with missing values and is known for its automated user interface. TPOT uses genetic programming. It builds pipelines over multiple generations, which can be time consuming. It has built-in pre-processing for missing values, but the genetic programming can require fine-tuning from users.

@@

\subsection{Training}

Insert Table Here
Table 2: Meta-features extracted in simplifcation of the Kaggle dataset-optimal model dataset

We used a Kaggle API library to search Kaggle using 20 medically-relevant key words. We limited our search to tabular datasets and obtained 300 viable datasets for training. We ran the Kaggle datasets through a rudimentary framework, pairing them with ML algorithms predicted to perform most optimally on that data. These ML algorithms were untrained and were selected from a pool of 30 candidates. The resulting 2x300 matrix then went through manuel meta-feature extraction (See Table 2). This new matrix was then used to train an XGBoost algorithm using a rank-based prediction system.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{geometry}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}

@@

\textbf{Ease of Use} & Easy-to-use GUI and API & Python library, requires coding experience & Very user-friendly web interface, detailed API & Simple API & User-friendly API with minimal coding & Python library with fairly intuitive API\\
\hline
\textbf{Supported Algorithms} & Linear, Tree, DNN, Clustering, etc. & Genetic Programming & Linear, Tree, LightGBM, CatBoost, Clustering, etc. (no DNN) & Linear, Tree, Clustering - More limited and lightweight & Linear, LightGBM, Catboost, etc. & Linear, tree, SVM, etc. \\
\hline
\textbf{Feature Engineering} & Basic & Basic & Advanced - includes feature interactions & Limited & Advanced (automatic) & Basic \\
\hline
\textbf{Time Management} & Time Constraints are Settable & Generations and population size & Time constraints are Settable & Optimized for quick results & Settable time constraints & Settable time constraints\\
\hline
\textbf{Ensemble Methods} & Yes (stacking + blending) & Yes (stacking) & Yes (stacking + blending) & No & Yes (stacking) & Yes (stacking) \\
\hline
\textbf{Hyperparameter Tuning} & Automated & Genetic algorithm & Automated and customizable & Automated and efficient tuning & Automated & Automated \\
\hline
\textbf{Model Interpretability} & SHAP, LIME, PDPs, MOJO & POJO, etc. & Basic & SHAP, LIME, PDPs, etc. & Basic (permutation importance) & SHAP, PDPs, etc. Basic (feature importance) \\
\hline
\textbf{Scalability} & Good (distributed computing) & Limited & Good & Good (resource-efficient) & Good (distributing computing) & Moderate \\
\hline
\textbf{Customization} & High & Moderate & High & Low-Moderate & Moderate & Moderate \\
\hline

@@

Taking a closer look at four AutoML standards, we can observe pros and cons in their theoretical application to medicine and research. FLAML uses blend search hyper-parameter optimization and focuses on lightweight models, making it quick and applicable to repetitive tasks in the medical sphere. It supports imputation techniques for missing values with user specification and has an intuitive API, making it moderately user-friendly for minimally trained healthcare workers. H2O has a wider variety of supported algorithms that may or may not be appropriate for medical diagnosis, making it slower than FLAML on Mean. In one trial, more than half of FLAML's performances in one minute were better than or equal to H20's performances in one hour. H2O has automatic and multi-faceted methods of dealing with missing values and flexible interfaces in R and Python, making it intuitive for users. MLJAR uses more advanced algorithms such as light gradient boosting and neural networks. It automatically deals with missing values and is known for its automated user interface. TPOT uses genetic programming. It builds pipelines over multiple generations, which can be time consuming. It has built-in pre-processing for missing values, but the genetic programming can require fine-tuning from users.

@@

\subsection{Overview}
When considering a specific dataset, various aspects regarding the distribution and relationship within the data can give us clues as to which models are likely to perform best. For example, linear models tend to work better on datasets where there is a linear relationship between features. As a result, we can develop a boosting model to predict the best models based off extracted attributes from a dataset.

\begin{table}
  \caption{Model Options}
  \label{model-options-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    Attribute & Model & Purpose                         \\
    \midrule
    \# of Rows & $|\mathbf{A}|_r$ & Data Size \\
    \hline
    \# of Columns & $|\mathbf{A}|_c$ & Data Size \\
    \hline
    Target Distribution &
    $\frac{\max(\text{count}_1, \text{count}_0)}{|\mathbf{A}|_r}$
    & Data Type \\
    \hline
    Numerical Columns &
    $\frac{|\mathbf{A}_{num}|}{|\mathbf{A}|_{c}}$
    & Data Type \\
    \hline
    Binary Categorical Columns &
    $\frac{|\mathbf{A}_{bin}|}{|\mathbf{A}|_{c}}$
    & Data Type \\
    \hline
    Mean \# of Distinct Values  & $\mu \left\{\text{\# of distinct values in column}\right\}$ & Data Type \\ 
    \hline
    Mean IQR  & $\mu \left\{\text{IQR of column}\right\}$ & Data Distribution \\ 
    \hline
    Standard Deviation of IQR  & $\sigma \left\{\text{IQR of column}\right\}$ & Data Distribution \\
    \hline
    Mean Q1 & $\mu \left\{\text{Q1 of column}\right\}$ & Data Distribution \\
    \hline
    Standard Deviation of Q1  & $\sigma \left\{\text{Q1 of column}\right\}$ & Data Distribution \\
    \hline
    Mean Q3  & $\mu \left\{\text{Q3 of column}\right\}$ & Data Distribution \\
    \hline
    Standard Deviation of Q3  & $\sigma \left\{\text{Q3 of column}\right\}$ & Data Distribution \\
    \hline
    Mean Percent of Z-Score Outliers  & 
    $\mu \left\{ \frac{\text{\# of Z-Score Outliers}}{|\mathbf{A}|_r} \right\}$ & Outlier Data Points \\
    \hline
    Standard Deviation of Percent of Z-Score Outliers  &
    $\sigma \left\{ \frac{\text{\# of Z-Score Outliers}}{|\mathbf{A}|_r} \right\}$ & Outlier Data Points \\
    \hline
    Mean Correlation& $\mu \left\{\text{Correlation Matrix}\right\}$ & Data linearity \\
    \hline
    Standard Deviation of Correlation & $\sigma \left\{\text{Correlation Matrix}\right\}$ & Data Linearity \\
    \bottomrule
  \end{tabular}
\end{table}
